An Embedding Framework for Consistent Polyhedral Surrogates

### 2021-04-03 poly-embed-journal/poly-embed-journal.tex
adding construction of S

In this section, we establish a tight relationship between the technique of embedding and the use of polyhedral (piecewise-linear convex) surrogate losses.
We defer the question of when such surrogates are consistent to \S~\ref{sec:calibration}.

@@

@@

\raf{We should probably add the type of $\ell$ and $L$ in each lemma/prop/thm statement, or add something like: ``throughout this section, unless specified otherwise, we use $\ell:\R\to\reals^\Y$ and $L:\reals^d\to\reals^\Y$''.}

\begin{proposition}\label{prop:embed-bayes-risks}
  Let discrete loss $\ell$ be given.
  Then $L$ embeds $\ell$ if and only if $\risk{L}=\risk{\ell}$.
  % Let discrete loss $\ell:\R\to\reals^Y$ be given.
  % Then $L:\reals^d\to\reals^\Y$ embeds $\ell$ if and only if $\risk{L}=\risk{\ell}$.
  % A loss $L$ embeds a discrete loss $\ell$ if and only if $\risk{L}=\risk{\ell}$.
\end{proposition}
\begin{proof}
  \raft{4/2/2021 first part of the proof should be done now.}
  Define $\Gamma = \prop{L}$ and $\gamma = \prop{\ell}$.
  Suppose $L$ embeds $\ell$, so we have some $\Sc\subseteq \R$ which is representative for $\ell$ and an embedding $\varphi:S\to\reals^d$.
  By Lemma~\ref{lem:loss-restrict}, we have $\risk{\ell} = \risk{\ell|_{\Sc}}$.
  Since $\Sc$ is representative for $\ell$, for all $p\in\simplex$ we have some $r \in \Sc \cap \gamma(p)$ and thus we have $\varphi(r) \in \U \cap \Gamma(p)$.
  In particular, $\U \cap \Gamma(p)\neq \emptyset$ for all $p\in\simplex$, so $\U$ is representative for $L$.
  \raft{Commented out: the symmetric argument that representation goes both ways under an embedding.  I'm thinking we should expand on this perspective in section 5, maybe introduce a notion of equivalence between any two losses (discrete or otherwise), and define trim in that context.}
  % Taking $\U := \varphi(\Sc)$, we have that $\ell$ if and only if $\U$ is representative of $L$, as $\gamma(p) \cap \Sc \neq \emptyset \iff \exists r \in \Sc$ such that $r \in \gamma(p) \iff \exists r$ such that $\varphi(r) \in \Gamma(p) \cap \U \iff \Gamma(p) \cap \U \neq \emptyset$.
%  To see that $\gamma'(p) \neq \emptyset$ for all $p\in\simplex$, note that by the definition of $\gamma$ as the property elicited by $\ell$ we have some $r \in \gamma(p)$, and by the embedding condition~\eqref{eq:embed-loss}, $\varphi(r) \in \Gamma(p)$.
  By Lemma~\ref{lem:loss-restrict}, we have $\risk{L} = \risk{L|_{\U}}$.
  As $L(\varphi(\cdot)) = \ell(\cdot)$ by embedding condition (i), for all $p\in\simplex$ we have
  \begin{equation*}
    \risk{\ell}(p) = \risk{\ell|_\Sc}(p) = \min_{r \in \Sc}\inprod{p}{\ell(r)} = \min_{r \in \Sc}\inprod{p}{L(\varphi(r))} = \min_{u \in \U}\inprod{p}{L(u)} = \risk{L|_\U}(p) = \risk{L}(p)~.
  \end{equation*}

    \btw{RF: This is on the informal side, mostly because of the ``working in the affine hull'' bit.  It can all be formalized though, probably should be for the journal version: just take an appropriate linear projection of $\reals^n$ to $\reals^{n-1}$, and work with the projected versions of everything, and then project back.}
  For the reverse implication, assume that $\risk{L} = \risk{\ell}$.
  In what follows, we implicitly work in the affine hull of $\simplex$ by taking an appropriate invertible linear projection to an affine subspace $A : \reals^n_+ \to \reals^{n-1}_+$, so that interiors are well-defined, and $\risk{\ell}$ may be differentiable on the interior of $\simplex$.
  (For intuition, one can consider the relative interiors in $\reals^\Y_+$, but we need to consider interiors in order to apply previous results.)
  Since $\ell$ is discrete, $-\risk{\ell}$ is polyhedral as the pointwise maximum of a finite set of linear functions.
  \raft{I believe we can remove mentions of power diagrams here; let's give it a try.  I don't think we need to project yet, given what we say next.}
  The projection of its epigraph $E_{\ell}$ onto $\simplex$ forms a power diagram by Theorem~\ref{thm:aurenhammer}, whose cells are full-dimensional (in $n-1$ dimensions) and correspond to the level sets $\gamma_r$ of $\gamma = \prop{\ell}$. \jessiet{Don't need Theorem 1; can use Lemma~\ref{lem:finite-full-dim} instead. -21 Mar 21}

@@

\proposedadd{For the reverse implication, assume $\risk{L} = \risk{\ell}$ are polyhedral functions.  Since $\risk{\ell}$ is polyhedral, the epigraph of its negation, $E_\ell$, has a finite set of facets, which by definition have well-defined interiors.  For each facet, there must then be some distribution $p_r$ and report $r \in \R$ such that the hyperplane $p_r \mapsto (p_r, \inprod{p_r}{-\ell(r)})$ uniquely supports $E_\ell$.  Form $\Sc$ from the set of such reports $r$, iterating over each facet.  We have $\Sc$ representative of $\ell$ since the union of projected epigraph facets equals the simplex, and these facets and their projections are closed. Therefore, every $p \in \simplex$ is contained in some facet of the epigraph, and thus there is some $r \in \Sc$ so that $r \in \gamma(p)$.
As the risks are equal, the same is true for $E_L$, e.g., for each facet, there is some $u_r$ so that $p_r \mapsto (p_r, \inprod{p_r}{L(u_r)})$, and one can construct the embedding $\varphi: r \mapsto u_r$ for each $r \in \Sc$.}

@@

\begin{lemma}
	\label{lem:poly-loss-poly-risk}
	If $L$ is polyhedral, $\risk{L}$ is polyhedral.
\end{lemma}
\begin{proof}
	Let $L:\reals^d\to\reals_+^\Y$ be a polyhedral loss, and $\Gamma = \prop{L}$.
	By Lemma~\ref{lem:polyhedral-range-gamma}, $\U = \Gamma(\simplex)$ is finite.
	For each $U\in \U$, select $u_U \in U$, and let $\Sc' = \{u_U : U \in\U\}$.
	$\Sc$ is representative of $L$\raft{let's use ``representative for'' to be consistent; can always find/replace to change later \#burn}, so Lemma~\ref{lem:loss-restrict} gives us $\risk{L} = \risk{L|_{\Sc}}$, which is polyhedral as $\Sc$ is finite.
\end{proof}

>>>>

In this section, we establish a tight relationship between the technique of embedding and the use of polyhedral (piecewise-linear convex) surrogate losses.
We defer the question of when such surrogates are consistent to \S~\ref{sec:calibration}. 
Throughout this section, unless specified otherwise, we use $\ell:\R\to\reals^\Y$ and $L:\reals^d\to\reals^\Y$

@@

\jessie{Adding new statements}
\begin{definition}
	Suppose a generic loss $L : \R \to \reals^\Y$ is given with polyhedral risk $-\risk{L}$.
	Construct the set $\Sc(L)$ as follows:
	Each facet $F$ of the epigraph of negative risk $E_L$ must be full dimensional by definition of a facet.
	Take $p \in \inter{F}$, and observe that it must by uniquely supported by some hyperplane $p \mapsto (p, L(u))$ for some $u \in \R$.
	Add $u$ to $\Sc(L)$, and repeat, iterating over each of $E_L$'s finitely many facets.
\end{definition}

\begin{proposition}
	If loss $L: \R \to \reals^\Y_+$ has polyhedral risk, $\Sc(L)$ is a minimal representative set for $L$.
\end{proposition}
\begin{proof}\jessiet{I don't like this proof, but meh.}
	If $\Sc(L)$ was not a minimal set, then we could remove some $s$ from $\Sc$ such that $\Sc' := \Sc \setminus \{s\}$ is also a representative set for $L$.
	By construction, we have some distribution $p$ such that the hyperplane $p \mapsto (p, L(s))$ uniquely supports $E_L$.
	We particularly have $\Gamma(p) \cap \Sc' = \emptyset$, since any other element $s'$ of $\Sc'$ being in $\Gamma(p)$ would imply $p \mapsto (p, L(s'))$ also supports $E_L$, but this hyperplane must be unique, and therefore either $s$ or $s'$ would not be in $\Sc$.
\end{proof}

\jessie{End additions}

@@

\begin{proposition}\label{prop:embed-bayes-risks}
  Let discrete loss $\ell$ be given.
  Then $L$ embeds $\ell$ if and only if $\risk{L}=\risk{\ell}$.
  % Let discrete loss $\ell:\R\to\reals^Y$ be given.
  % Then $L:\reals^d\to\reals^\Y$ embeds $\ell$ if and only if $\risk{L}=\risk{\ell}$.
  % A loss $L$ embeds a discrete loss $\ell$ if and only if $\risk{L}=\risk{\ell}$.
\end{proposition}
\begin{proof}
  \raft{4/2/2021 first part of the proof should be done now.}
  Define $\Gamma = \prop{L}$ and $\gamma = \prop{\ell}$.
  Suppose $L$ embeds $\ell$, so we have some $\Sc\subseteq \R$ which is representative for $\ell$ and an embedding $\varphi:\Sc\to\reals^d$.
  By Lemma~\ref{lem:loss-restrict}, we have $\risk{\ell} = \risk{\ell|_{\Sc}}$.
  Since $\Sc$ is representative for $\ell$, for all $p\in\simplex$ we have some $r \in \Sc \cap \gamma(p)$ and thus we have $\varphi(r) \in \U \cap \Gamma(p)$.
  In particular, $\U \cap \Gamma(p)\neq \emptyset$ for all $p\in\simplex$, so $\U$ is representative for $L$.
  \raft{Commented out: the symmetric argument that representation goes both ways under an embedding.  I'm thinking we should expand on this perspective in section 5, maybe introduce a notion of equivalence between any two losses (discrete or otherwise), and define trim in that context.}
  % Taking $\U := \varphi(\Sc)$, we have that $\ell$ if and only if $\U$ is representative of $L$, as $\gamma(p) \cap \Sc \neq \emptyset \iff \exists r \in \Sc$ such that $r \in \gamma(p) \iff \exists r$ such that $\varphi(r) \in \Gamma(p) \cap \U \iff \Gamma(p) \cap \U \neq \emptyset$.
%  To see that $\gamma'(p) \neq \emptyset$ for all $p\in\simplex$, note that by the definition of $\gamma$ as the property elicited by $\ell$ we have some $r \in \gamma(p)$, and by the embedding condition~\eqref{eq:embed-loss}, $\varphi(r) \in \Gamma(p)$.
  By Lemma~\ref{lem:loss-restrict}, we have $\risk{L} = \risk{L|_{\U}}$.
  As $L(\varphi(\cdot)) = \ell(\cdot)$ by embedding condition (i), for all $p\in\simplex$ we have
  \begin{equation*}
    \risk{\ell}(p) = \risk{\ell|_\Sc}(p) = \min_{r \in \Sc}\inprod{p}{\ell(r)} = \min_{r \in \Sc}\inprod{p}{L(\varphi(r))} = \min_{u \in \U}\inprod{p}{L(u)} = \risk{L|_\U}(p) = \risk{L}(p)~.
  \end{equation*}

    %\btw{RF: This is on the informal side, mostly because of the ``working in the affine hull'' bit.  It can all be formalized though, probably should be for the journal version: just take an appropriate linear projection of $\reals^n$ to $\reals^{n-1}$, and work with the projected versions of everything, and then project back.}
  For the reverse implication, assume that $\risk{L} = \risk{\ell}$.
  In what follows, we implicitly work in the affine hull of $\simplex$ by taking an appropriate invertible linear projection to an affine subspace $A : \reals^n_+ \to \reals^{n-1}_+$, so that interiors are well-defined, and $\risk{\ell}$ may be differentiable on the interior of $\simplex$.
  (For intuition, one can consider the relative interiors in $\reals^\Y_+$, but we need to consider interiors in order to apply previous results.)
  Since $\ell$ is discrete, $-\risk{\ell}$ is polyhedral as the pointwise maximum of a finite set of linear functions.
  \raft{I believe we can remove mentions of power diagrams here; let's give it a try.  I don't think we need to project yet, given what we say next.}
  The projection of its epigraph $E_{\ell}$ onto $\simplex$ forms a power diagram by Theorem~\ref{thm:aurenhammer}, whose cells are full-dimensional (in $n-1$ dimensions) and correspond to the level sets $\gamma_r$ of $\gamma = \prop{\ell}$. \jessiet{Don't need Theorem 1; can use Lemma~\ref{lem:finite-full-dim} instead. -21 Mar 21}

@@

\proposedadd{For the reverse implication, assume $\risk{L} = \risk{\ell}$ are polyhedral functions.  Since $\risk{\ell}$ is polyhedral, the epigraph of its negation, $E_\ell$, has a finite set of facets, which by definition have well-defined interiors.  For each facet, there must then be some distribution $p_r$ and report $r \in \R$ such that the hyperplane $p_r \mapsto (p_r, \inprod{p_r}{-\ell(r)})$ uniquely supports $E_\ell$.  Form $\Sc$ from the set of such reports $r$, iterating over each facet.  We have $\Sc$ representative for $\ell$ since the union of projected epigraph facets equals the simplex, and these facets and their projections are closed. Therefore, every $p \in \simplex$ is contained in some facet of the epigraph, and thus there is some $r \in \Sc$ so that $r \in \gamma(p)$.
As the risks are equal, the same is true for $E_L$, e.g., for each facet, there is some $u_r$ so that $p_r \mapsto (p_r, \inprod{p_r}{L(u_r)})$, and one can construct the embedding $\varphi: r \mapsto u_r$ for each $r \in \Sc$.}

@@

\begin{lemma}
	\label{lem:poly-loss-poly-risk}
	If $L$ is polyhedral, $\risk{L}$ is polyhedral.
\end{lemma}
\begin{proof}
	Let $L:\reals^d\to\reals_+^\Y$ be a polyhedral loss, and $\Gamma = \prop{L}$.
	By Lemma~\ref{lem:polyhedral-range-gamma}, $\U = \Gamma(\simplex)$ is finite.
	For each $U\in \U$, select $u_U \in U$, and let $\Sc' = \{u_U : U \in\U\}$.
	$\Sc$ is representative for $L$, so Lemma~\ref{lem:loss-restrict} gives us $\risk{L} = \risk{L|_{\Sc}}$, which is polyhedral as $\Sc$ is finite.
\end{proof}
